\lightkurve: an open source Python package for NASA's \kepler, \ktwo~ and \tess~ data analysis

### 2018-04-04 paper/ms.tex
Updated structure

\newcommand{\ktwo}{{\it K$\mathit{2}$}}
\newcommand{\tess}{{\it TESS}}
\newcommand{\kepler}{{\it Kepler}}
\newcommand{\lightkurve}{\texttt{lightkurve}}
\newcommand{\LightCurve}{\texttt{LightCurve}}
\newcommand{\KeplerLightCurve}{\texttt{KeplerLightCurve}}
\newcommand{\numpy}{\texttt{numpy}}
\newcommand{\code}{\texttt}
\newcommand{\clh}[1]{\textcolor{red}{ \textbf{***CLH: #1 ***}}}
\newcommand{\ze}[1]{\textcolor{blue}{ \textbf{***CLH: #1 ***}}}

@@

\begin{abstract}

  \begin{itemize}
  \item{We present \lightkurve, an open source package for analysis of flux time series using Python.}
  \item{It is a multipurpose, general tool to make producing time-series photometry from raw pixel data easier and more reproducible between different teams.}
  \item{\lightkurve~ is designed to interface seemlessly with data from NASA's \kepler, \ktwo~ and \tess~ missions.}
  \item{Using these tools it is simple to create corrected time-series photometry from raw pixel data from any of these missions.}
  \item{Easy to interact with raw data products and produce fantastic light curves}
  \item{Open source, modable}
  \item{Supremely reproducible.}
  \item{Go from raw data to planet in 5 lines of code.}
  \end{itemize}

@@

\paragraph{What tools are currently available?}
\begin{itemize}
\item PyKE
\item \clh{There must be others?}
\item However, no one package provide a simple, open source framework for manipulating time-series data that is general purpose.
\end{itemize}

@@

\paragraph{What is the selling point of lightkurve}
\begin{itemize}
\item There are two sides to the \lightkurve package. Firstly, \lightkurve can be used as an extraction package for creating time-series photometry from astronomical images such as \kepler Target Pixel Files (TPFs) or \tess Full Frame Images (FFIs). This includes simple aperture photometry, PSF photometry and centroiding.
\item Secondly, \lightkurve can be used for analysis of time-series photometry. This includes motion detrending, CBV corrections, outlier rejection and period folding.
\item Together these two sides can be combined to convert raw data from \kepler, \ktwo~ and \tess~ to cleaned light curves of exoplanet candidates, supernovae and extra-galactic objects.
\end{itemize}

\paragraph{Future resources?}
\begin{itemize}
\item This is \lightkurve 1.0
\item Anticipate adding new features
\item Easily extendible for users to add in new features
\end{itemize}

@@

\clh{This needs to VERY clearly show the difference between a TPF, LC and LCF.}
Maybe include a class diagram here, or a table of the modules or classes.

Maybe include a table or graph listing the key functions of the key classes,
e.g.
Lightcurve: cdpp, fold, stitch, plot
TargetPixelFile: tolightcurve, plot

@@

\section{Examples of key features}

\subsection{Working with lightcurves}

flatten, fold, finding planets, from archive...

\subsection{Making your own lightcurves from target pixel files}

explain TPFs, aperture photometry, and PSF photometry

\subsection{Removing systematics from lightcurves}

@@

@@

\section{Lightcurve Basics}

        \begin{table}[!htb]
            \centering
            \caption{A subset of methods provided by the \LightCurve~class}
            \begin{tabular}{cp{6.5cm}}
                \hline
                \textbf{Method} & \textbf{Short description} \\
                \hline
                \texttt{stitch} & appends the attributes \texttt{flux},
                \texttt{time}, and \texttt{flux\_err} of other given
                \texttt{LighCurve} objects.\\
                \texttt{flatten} & applies a Savitzky-Golay filter to capture
                low frequency flux variations which can be then removed in order
                to aid transit detection algorithms.\\
                \texttt{fold} & folds a lightcurve at a given period and phase.\\
                \texttt{bin} &  bins a lightcurve using a block mean or median.\\
                \texttt{cdpp} &  computes the Combined Differential Photometric
                Precision (CDPP) metric, which is a proxy for the amount of
                scatter in the lightcurve signal. \\
                \texttt{plot} & displays a lightcurve.
            \end{tabular}
            \label{tab:methods}
        \end{table}

\begin{figure}
\plotone{figs/fold-lc.pdf}
\caption{Folded lightcurve of target \texttt{KIC011904151} quarter 3, showing the
            transit signal of Kepler-10b.
\label{fig:fold-method}}
\end{figure}

\section{Target Pixel File Basics}
    \subsection{The \texttt{KeplerTargetPixelFile} class}
        A \texttt{KeplerTargetPixelFile} object can be instantiated
        by passing a \texttt{path} (URL or local) of a target pixel file.
        Optionally, the user can elect to throw away frames that contain
        a specific flag by using the \texttt{quality\_bitmask} argument.

        For instance, the method \texttt{plot} can be used to visualize a
        given frame, which are depicted in Fig.~\ref{fig:plot-method}.
\begin{verbatim}
>>> import numpy as np
>>> from lightkurve import KeplerTargetPixelFile
>>> tpf = KeplerTargetPixelFile("kplr008462852-2011073133259_lpd-targ.fits")
>>> tpf.plot()
>>> tpf.plot(aperture_mask=tpf.flux[0] > np.nanmean(tpf.flux[0]))
\end{verbatim}

\section{Tools}

\subsection{Cotrending basis vectors}

Cotrending basis vectors (CBVs) can remove global correlated
systematics present in a given channel \cite{smith2012}.

@@

\begin{figure}[!htb]
    \centering
    \plotone{figs/cbv.pdf}
    \caption{CBV correction applied on \texttt{KOI 8462852}}
    \label{fig:cbv-correction}
\end{figure}

Improperly tuning the number of CBVs can cause over-/under-fitting. One
way to identify a reasonable number of CBVs is to perform a grid search
as shown in Fig~(\ref{fig:cbv-grid-search}).

Fig.~\ref{fig:cbv-correction-k2} shows the detrended lightcurve after estimating
the first nine coefficients for CBV correciton on \ktwo target
 \texttt{EPIC 201543306}. The selection of the number of CBVs is set by inspecting the
grid search curve can be set through model comparison heuristics like AIC, BIC,
or cross-validation \cite{ivezi2014}.

\begin{figure}
    \centering
    \plotone{figs/cbv-k2.pdf}
    \caption{CBV correction applied on \texttt{EPIC 201543306}.
    \label{fig:cbv-correction-k2}}
\end{figure}

\subsubsection{Number of CBVs}
The number of CBVs will directly contribute to overfitting effects. One
way to identify a reasonable number of CBVs is to perform a grid search
as suggested in Fig~(\ref{fig:cbv-grid-search}), which shows the cost
function as a function of the number of CBVs. Usually, as the number of
CBVs increases, the value of the cost function decreases. And therefore,
the user should empirically choose a number of CBVs which does not
remove the astrophysical signal of interest [add reference].

\begin{figure}[!htb]
    \centering
    \plotone{figs/cbv-grid-search.pdf}
    \caption{Grid search on the number of CBVs}
    \label{fig:cbv-grid-search}
\end{figure}

% MGS note: Wait--- won't the cost function decrease monotonically for CBVs?
% Do you use AIC or BIC or cross-validation?
An objective way of selecting the number of CBVs is to use Bayes' factors
[add reference]. In the Bayes' factor setting, the selected number of
CBVs is the one that provide the least gain in posterior probability, i.e.,
for all ordered pairs of CBVs, the Bayes factor selects $n^{\star}$ number of CBVs
as follows
\begin{align}
    n^{\star} = \argmin_{n} \dfrac{p_{n+1}}{p_n},
\end{align}
in which $p_n$ is the posterior probability evaluated at the Maximum A Posteriori
Estimator (MAP) obtained using $n$ CBVs.

@@

\begin{figure}[!htb]
    \centering
    \plotone{figs/cbv-bayes-factor.pdf}
    \caption{Bayes factors as a function of the number of CBVs for \texttt{EPIC 201543306}}
    \label{fig:cbv-grid-search}
\end{figure}

\subsection{Point spread function photometry}

Point Spread Function (PSF) photometry is the de facto technique to
process crowded-field images~\cite{stetson1987, heasley1999}. In context of Kepler
and K2 missions, Libralato \textit{et al}~\cite{libralato2016} have shown...

The underlying principle of PSF photometry consists in modelling a given crowded
image as a linear combination of individual PSFs and possibly a background model.
On the PSF model itself, it is commonly assumed that the flux
at an arbitrary pixel position increases linearly with the integrated
flux~\cite{stetson1987, heasley1999}.

\lightkurve ~contains routines to perform PSF photometry in TPFs
which are implemented in the \texttt{psf} module.

@@

\begin{verbatim}
>>> from lightkurve import KeplerTargetPixelFile
>>> from lightkurve.psf import PRFPhotometry, SceneModel
>>> from oktopus import UniformPrior
>>> tpf = KeplerTargetPixelFile("ktwo246199087-c12_lpd-targ.fits.gz")
>>> prf = tpf.get_prf_model()
>>> prior = UniformPrior(lb=[4e3, 990, 25, 1], ub=[2e4, 996, 30, 2e3])
>>> scene = SceneModel(prf=[prfs])
>>> phot = PRFPhotometry(scene_model=scene, prior=prior)
>>> results = phot.fit(tpf.flux + tpf.flux_bkg)
\end{verbatim}

@@

\subsection{Motion-dependent Correlated Noise}
\label{subsection:motion}

Spacecraft-induced correlated noise remains one of the greatest hurdles to
analyzing K2 lightcurves. Many algorithms have been developed to mitigate
motion-dependent artifacts~\cite{vanderburg14}[CITE K2SC and EVEREST].
In \lightkurve, we implement an algorithm based off of the self-flat-field
(SFF) presented in~\cite{vanderburg14}.

Basically, SFF works by decorrelating the simple aperture flux
against the information on the spacecraft motion, obtained by computing the
arclength using the centroids of the target.

@@

We would like to express our gratitude...

@@

\appendix

\section{Appendix information}

>>>>

\newcommand{\ktwo}{{\it K$\mathit{2}$}}
\newcommand{\tess}{{\it TESS}}
\newcommand{\kepler}{{\it Kepler}}
\newcommand{\lightkurve}{\texttt{lightkurve}}
\newcommand{\LightCurve}{\texttt{LightCurve}}
\newcommand{\KeplerLightCurve}{\texttt{KeplerLightCurve}}
\newcommand{\numpy}{\texttt{numpy}}
\newcommand{\code}{\texttt}
\newcommand{\clh}[1]{\textcolor{red}{ \textbf{***CLH: #1 ***}}}
\newcommand{\ze}[1]{\textcolor{blue}{ \textbf{***CLH: #1 ***}}}
\newcommand{\gb}[1]{\textcolor{green}{ \textbf{***CLH: #1 ***}}}

@@

\begin{abstract}
\gb{Key messages are: WHY do users need to create their own products? (RB, asteroids, variable apertures, crowding etcetc). WHAT are we providing? (lightkurve, easy Kepler/K2/TESS access). HOW can they use it? (Beautiful examples)}
  \begin{itemize}
  \item{Performing custom photometry allows users to mitigate...}
  \item{However, with complex `big' data from \kepler and \ktwo...}
  \item{We present \lightkurve, an open source package for analysis of flux time series using Python.}
  \item{It is a multipurpose, general tool to make producing time-series photometry from raw pixel data easier and more reproducible between different teams.}
  \item{\lightkurve~ is designed to interface seemlessly with data from NASA's \kepler, \ktwo~ and \tess~ missions.}
  \item{Using these tools it is simple to create corrected time-series photometry from raw pixel data from any of these missions.}
  \item{Easy to interact with raw data products and produce fantastic light curves}
  \item{Open source, modable}
  \item{Supremely reproducible.}
  \item{Go from raw data to planet in 5 lines of code.}
  \item{Excellent learning tool}
  \end{itemize}

@@

\paragraph{What tools are currently available?}
\begin{itemize}
\item PyKE
\item \clh{Geert says he can do this para}
\item However, no one package provide a simple, open source framework for manipulating time-series data that is general purpose.
\end{itemize}

@@

\paragraph{What is the selling point of lightkurve}
\begin{itemize}
\item There are two sides to the \lightkurve package. Firstly, \lightkurve can be used as an extraction package for creating time-series photometry from astronomical images such as \kepler Target Pixel Files (TPFs) or \tess Full Frame Images (FFIs). This includes simple aperture photometry, PSF photometry and centroiding.
\item Secondly, \lightkurve can be used for analysis of time-series photometry. This includes motion detrending, CBV corrections, outlier rejection and period folding.
\item Together these two sides can be combined to convert raw data from \kepler, \ktwo~ and \tess~ to cleaned light curves of exoplanet candidates, supernovae and extra-galactic objects.
\item One flexible system for all optical photometry
\item Learning/teaching tool
\end{itemize}

\paragraph{Future resources?}
\begin{itemize}
\item This is \lightkurve 1.0
\item Anticipate adding new features
\item Easily extendible for users to add in new features
\item{There are already tutorials, which will be expanded}
\end{itemize}

@@

\clh{Ze can you do a beautiful diagram here?}

\subsection{The \texttt{KeplerTargetPixelFile} class}
        A \texttt{KeplerTargetPixelFile} object can be instantiated
        by passing a \texttt{path} (URL or local) of a target pixel file.
        Optionally, the user can elect to throw away frames that contain
        a specific flag by using the \texttt{quality\_bitmask} argument.

        For instance, the method \texttt{plot} can be used to visualize a
        given frame, which are depicted in Fig.~\ref{fig:plot-method}.

\begin{verbatim}
>>> import numpy as np
>>> from lightkurve import KeplerTargetPixelFile
>>> tpf = KeplerTargetPixelFile("kplr008462852-2011073133259_lpd-targ.fits")
>>> tpf.plot()
>>> tpf.plot(aperture_mask=tpf.flux[0] > np.nanmean(tpf.flux[0]))
\end{verbatim}

\clh{This table I believe needs a little updating? There can also be a second table for TPF}
\begin{table}[!htb]
    \centering
    \caption{A subset of methods provided by the \LightCurve~class}
    \begin{tabular}{cp{6.5cm}}
        \hline
        \textbf{Method} & \textbf{Short description} \\
        \hline
        \texttt{stitch} & appends the attributes \texttt{flux},
        \texttt{time}, and \texttt{flux\_err} of other given
        \texttt{LighCurve} objects.\\
        \texttt{flatten} & applies a Savitzky-Golay filter to capture
        low frequency flux variations which can be then removed in order
        to aid transit detection algorithms.\\
        \texttt{fold} & folds a lightcurve at a given period and phase.\\
        \texttt{bin} &  bins a lightcurve using a block mean or median.\\
        \texttt{cdpp} &  computes the Combined Differential Photometric
        Precision (CDPP) metric, which is a proxy for the amount of
        scatter in the lightcurve signal. \\
        \texttt{plot} & displays a lightcurve.
    \end{tabular}
    \label{tab:methods}
\end{table}

@@

\section{Common Use Cases}

\subsection{Creating Custom Light Curves}

    \subsubsection{Simple Aperture Photometry (SAP)}
    \begin{itemize}
        \item What is SAP?
        \item Why do SAP?s
        \item How do you do SAP?
    \end{itemize}

    \subsubsection{Point Spread Function (PSF) Photometry}
        %What is PSF?
        Point Spread Function (PSF) photometry is the de facto technique to
        process crowded-field images~\cite{stetson1987, heasley1999}. In context of Kepler
        and K2 missions, Libralato \textit{et al}~\cite{libralato2016} have shown...

        See a detailed explaination of PSF photometry in \ref{appendix:psf}
        %Why do PSF?
        The underlying principle of PSF photometry consists in modelling a given crowded
        image as a linear combination of individual PSFs and possibly a background model.
        On the PSF model itself, it is commonly assumed that the flux
        at an arbitrary pixel position increases linearly with the integrated
        flux~\cite{stetson1987, heasley1999}.

        %How do you do PSF?
        \lightkurve ~contains routines to perform PSF photometry in TPFs
        which are implemented in the \texttt{psf} module.

        \clh{If you want to include oktopus in this snippet you have to explain what it is to the reader:)}
        \begin{verbatim}
        >>> from lightkurve import KeplerTargetPixelFile
        >>> from lightkurve.psf import PRFPhotometry, SceneModel
        >>> from oktopus import UniformPrior
        >>> tpf = KeplerTargetPixelFile("ktwo246199087-c12_lpd-targ.fits.gz")
        >>> prf = tpf.get_prf_model()
        >>> prior = UniformPrior(lb=[4e3, 990, 25, 1], ub=[2e4, 996, 30, 2e3])
        >>> scene = SceneModel(prf=[prfs])
        >>> phot = PRFPhotometry(scene_model=scene, prior=prior)
        >>> results = phot.fit(tpf.flux + tpf.flux_bkg)
        \end{verbatim}

\subsection{Correcting Common Systematics}

    We provide tools to correct for two systematics that are common between targets on the same channel.
    We provide corrections using \emph{Cotrending Basis Vectors} (CBVs) which mitigate systematics due to e.g. spacecraft heating. (See Appendix \ref{appendix:CBV} for a detailed explanation of CBVs)
    We also provide a simple implementation of the \emph{Self Flat Fielding} (SFF) method to correct for spacecraft motion. (See Appendix \ref{appendix:motion} for a detailed explanation of SFF)

@@

    \subsubsection{Correcting Spacecraft Motion using Lightkurve}
        Spacecraft-induced correlated noise remains one of the greatest hurdles to
        analyzing K2 lightcurves. Many algorithms have been developed to mitigate
        motion-dependent artifacts~\cite{vanderburg14}[CITE K2SC and EVEREST].
        In \lightkurve, we implement an algorithm based off of the self-flat-field
        (SFF) presented in~\cite{vanderburg14}.

        SFF works by decorrelating the simple aperture flux
        against the information on the spacecraft motion, obtained by computing the
        arclength using the centroids of the target. (See Appendix \ref{appendix:motion} for a detailed explanation of SFF)

    \subsubsection{Correcting Common Systematics with CBVs using Lightkurve}

        \clh{Simple paragraph describing CBVs}
        Cotrending basis vectors (CBVs) can remove global correlated
        systematics present in a given channel \cite{smith2012}.

        \begin{figure}[!htb]
            \centering
            \plotone{figs/cbv.pdf}
            \caption{CBV correction applied on \texttt{KOI 8462852}
            \clh{This is great, really valuable plot. I would just change to have two lines with labels corrected and uncorrected.}}
            \label{fig:cbv-correction}
        \end{figure}

        Improperly tuning the number of CBVs can cause over-/under-fitting. One
        way to identify a reasonable number of CBVs is to perform a grid search
        as shown in Fig~(\ref{fig:cbv-grid-search}). The selection of the number of CBVs is set by inspecting the
        grid search curve can be set through model comparison heuristics like AIC, BIC,
        or cross-validation \cite{ivezi2014}.

\subsection{Recovering a planet signal}

%CLH: This might be unnecessary
\subsection{Light curves from large images}

\section{Example Code Snippets}

\subsection{Quickly recover a planet}
\begin{figure}
\plotone{figs/fold-lc.pdf}
\caption{Folded lightcurve of target \texttt{KIC011904151} quarter 3, showing the
            transit signal of Kepler-10b.
\label{fig:fold-method}}
\end{figure}

\subsection{Correct a supernova light curve}

\subsection{Background Subtraction for AGN}

@@

\appendix
\section{Cotrending basis vectors}
\label{appendix:cbvs}

@@

%CLH: I don't think we need this figure

%\begin{figure}
%    \centering
%    \plotone{figs/cbv-k2.pdf}
%    \caption{CBV correction applied on \texttt{EPIC 201543306}.
%    \label{fig:cbv-correction-k2}}
%\end{figure}

%I think this para will be moved in later versions, we don't need this figure,
The number of CBVs will directly contribute to overfitting effects. One
way to identify a reasonable number of CBVs is to perform a grid search
as suggested in Fig~(\ref{fig:cbv-grid-search}), which shows the cost
function as a function of the number of CBVs. Usually, as the number of
CBVs increases, the value of the cost function decreases. And therefore,
the user should empirically choose a number of CBVs which does not
remove the astrophysical signal of interest [add reference].

%\begin{figure}[!htb]
%    \centering
%    \plotone{figs/cbv-grid-search.pdf}
%    \caption{Grid search on the number of CBVs}
%    \label{fig:cbv-grid-search}
%\end{figure}

% MGS note: Wait--- won't the cost function decrease monotonically for CBVs?
% Do you use AIC or BIC or cross-validation?

%CLH: This is a little detailed for this paper, I might consider just leaving it as ``we use BIC''
An objective way of selecting the number of CBVs is to use Bayes' factors
[add reference]. In the Bayes' factor setting, the selected number of
CBVs is the one that provide the least gain in posterior probability, i.e.,
for all ordered pairs of CBVs, the Bayes factor selects $n^{\star}$ number of CBVs
as follows
\begin{align}
    n^{\star} = \argmin_{n} \dfrac{p_{n+1}}{p_n},
\end{align}
in which $p_n$ is the posterior probability evaluated at the Maximum A Posteriori
Estimator (MAP) obtained using $n$ CBVs.

@@

%\begin{figure}[!htb]
%    \centering
%    \plotone{figs/cbv-bayes-factor.pdf}
%    \caption{Bayes factors as a function of the number of CBVs for \texttt{EPIC 201543306}}
%    \label{fig:cbv-grid-search}
%\end{figure}

\section{Point spread function photometry}
\label{appendix:psf}

@@

@@

\section{Motion-dependent Correlated Noise}
\label{appendix:motion}

\clh{Quick discussion of the SFF algorithm}

@@

We would like to express our gratitude...
Funding sources

@@
